---
title: "COVID-19 Project"
output: 
  html_document:
    toc: true
    toc_float: true
---

## ***Gehad Gad***
## May 11th, 2020


# 1. Introduction:

Coronavirus Disease 2019 (COVID-19) is one of the coronavirus family that is usually common in animals. COVID-19 affects the respiratory and cause symptoms such as: cough, fever, shortness of breath, muscle aches, sore throat, unexplained loss of taste or smell, diarrhea and headache. COVID-19 can be severe, and some cases have caused death.
There are 4,013,728 confirmed cases and 278,993 Confirme deaths of COVID-19 

# 2. Project objectives
1.Determine the effective factors/features on spreading Coronavirus.
2.Detect the factor(s) that may lead to increase the fatality rate.
3.Build a classifier that can estimate the probability of being infected.


# 3. Methodolgy:


Data cleaning, prepration, statistical analysis metrics such as: mean, median, and correlation.The analysis include graphs; histograms, box plots, barplot, and time series plot to make better view.***Backward selection*** method to choose the most effective features indicated using the changes in model performance.***Random Forest*** algorithm to detect the most important features in relation with the infection and deaths. and finally is***Neural Network Model***.


# 4. Packages and libraries

```{r}
if(!require(dplyr)){
    install.packages("dplyr")
    library(dplyr)
}
if(!require(ggplot2)){
    install.packages("ggplot2")
    library(ggplot2)
}
if(!require(keras)){
    install.packages("keras")
    library(keras)
}
if(!require(fastDummies)){
    install.packages("fastDummies")
    library(fastDummies)
}
if(!require(caTools)){
    install.packages("caTools")
    library(caTools)
}
if(!require(caret)){
    install.packages("caret")
    library(caret)
}
if(!require(randomForest)){
    install.packages("randomForest")
    library(randomForest)
}
if(!require(e1071)){
    install.packages("e1071")
    library(e1071)
}
if(!require(rjson)){
    install.packages("rjson")
    library(rjson)
}
```

# 5. Research 1: Data analysis

## Datasets used :
1.COVID-19 global report
2. Countries’ space or size and GDP dataset
3. Countries detailed reports for literacy, 4. foods dataset
5. Countries detailed reports for production, 6. healthcare quality care quality, education dataset 



```{r}

#Import the data
global_data <- read.csv('train.csv')

#Display the first few fows.
head(global_data)
```

# The ID column is not an effective feature as it represents the observation index so it must be deleted

```{r}
#Remove the ID column
global_data = global_data[-c(1)]
```

# Examin the structure of the data

```{r}
str(global_data)
```

# As we see both countries and state have a certain values so , they must be categorical data


```{r}
#The date in this data is defined as factor and to do time series analysis, we need to convert it into date type

global_data$Date = as.character(global_data$Date)
global_data$Date = as.Date(global_data$Date, format = "%Y-%m-%d")


str (global_data)
```



```{r}
# Analyze the data using decribtive statistics. we will start by analyzing the continues data

confirmed_countries = aggregate(global_data$ConfirmedCases,list(global_data$Country.Region),sum)

print(confirmed_countries)
```


```{r}
#Gather all the fatalities in each country

death_countries = aggregate(global_data$Fatalities,list(global_data$Country.Region),sum)

print(death_countries)
```


```{r}
Countries_means = death_countries

# Renaming the x column to mean_deaths

colnames(Countries_means)[2] <-"Fatalities"

# Adding the mean infection to the data frame called (Countries_means). 

Countries_means['ConfirmedCases'] = confirmed_countries$x

head(Countries_means)
```

```{r}
#import the geo info for the countries
#data source:
#geo = read.csv('Counties geo.csv')
colnames(geo)[2] <-"Area"
head(geo)
```

```{r}
# Match the geographical data with the first data (global_data)

global_data$Area <- geo$Area[match(global_data$Country.Region, geo$Country)]

global_data$population <- geo$Population[match(global_data$Country.Region, geo$Country)]

global_data$Continent <- geo$Continent[match(global_data$Country.Region, geo$Country)]

print (global_data)
```


```{r}
#population per kilometer 

global_data$pop_per_k = global_data$population / global_data$Area

head(global_data)
```

```{r}
#Gather the of the population per kilometer for each country.

mean_numbers_per_k = aggregate(global_data$pop_per_k,list(global_data$Country.Region),mean)

print(mean_numbers_per_k)
```

```{r}
#Merge the population ratio to our countries_means data. 
Countries_means['population_ration'] = mean_numbers_per_k$x

#rename the cloumn Group.1 to Country

colnames(Countries_means)[1] <-"Country"

#Display the head
print(Countries_means)

```



```{r}
# Remove the outliers

#Make a copy of our data (Countries_means)
country = Countries_means

#Search for outliers in population_ration.
outliers <- boxplot(Countries_means$population_ration, plot=FALSE)$out
Countries_means<- Countries_means[-which(Countries_means$population_ration %in% outliers),]

#Check for outliers in Fatalities
outliers <- boxplot(Countries_means$Fatalities, plot=FALSE)$out
Countries_means<- Countries_means[-which(Countries_means$Fatalities %in% outliers),]

#Check for outliers in Confirmed cases
outliers <- boxplot(Countries_means$ConfirmedCases, plot=FALSE)$out
Countries_means<- Countries_means[-which(Countries_means$ConfirmedCases %in% outliers),]


```


```{r}
hist(Countries_means$Fatalities)
```

The histogram plot showa that the most frequent deaths is between 0 and 5 deaths. We can tell that the data of the Fatalities cases is not normally distributed, which will help us in our descriptive analysis


```{r}
hist(Countries_means$ConfirmedCases )

```


The histogram plot that the most frequent infections is between 0 and 500 cases. We can tell that the data of the Confirmed cases is not normally distributed, which will help us in our descriptive analysis

```{r}
hist(Countries_means$population_ration )

```

The histogram plot shows the most frequent population per kilometer is between 0 and 100. We can tell that the data of the Fatalities cases is not normally distributed, which will help us in our descriptive analysis

```{r}
#Find the confirmed and deaths cases by continents.
Continents_casses_sum = aggregate(global_data$ConfirmedCases,list(global_data$Continent),sum)
Continents_deaths_sum = aggregate(global_data$Fatalities,list(global_data$Continent),sum)

#Combine the two columns in a new data frame called (Continents_means)
Continents_means = Continents_casses_sum
Continents_means['Fatalities'] = Continents_deaths_sum$x

#Renaming the columns names
colnames(Continents_means)[2] <-"Confirmed"
colnames(Continents_means)[1] <-"continent"

#Display the head
head(Continents_means)

```


```{r}

#Get the difference between the Confirmed and Fatalities.

Continents_means$Difference = Continents_means$Confirmed / Continents_means$Fatalities


head(Continents_means)


```


```{r}

ggplot(data=Continents_means, aes(x=continent, y=Fatalities)) +
  geom_bar(stat="identity")
```

we have indication that Asia is the highest country with Fatalities, Europe is in the second place

```{r}
ggplot(data=Continents_means, aes(x=continent, y=Confirmed)) +
  geom_bar(stat="identity")
```

we have indication that Asia is the highest country with infected cases, Europe is in the second place.


```{r}
ggplot(data=Continents_means, aes(x=continent, y=Difference)) +
  geom_bar(stat="identity")
```


```{r}
#Time series of the confirmed cases
ggplot(global_data, aes(x=Date, y=ConfirmedCases)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("") 
```

```{r}
#Seprate the confirmed cases and fatalities cases
GlobalConfirmedCasses = aggregate(global_data$ConfirmedCases,list(global_data$Date),sum)
GlobalDeathCases = aggregate(global_data$Fatalities,list(global_data$Date),sum)

#Combine them in a new data frame and rename the columns

GlobalConfimred_Deaths = GlobalConfirmedCasses
GlobalConfimred_Deaths['Totaldeaths'] = GlobalDeathCases$x
colnames(GlobalConfimred_Deaths)[2] <-"casses"
colnames(GlobalConfimred_Deaths)[1] <-"Date"

#Display the head
head(GlobalConfimred_Deaths)
```

```{r}
ggplot(GlobalConfimred_Deaths, aes(x=Date, y=casses)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("") 
```



```{r}
ggplot(GlobalConfimred_Deaths, aes(x=Date, y=Totaldeaths)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("")
```

As we have another strong indication that as the time passes both death cases and infection cases increase all around the world. we can indicate a strong relationship between the infection and the deaths as both of the trends have the same shape 

```{r}

#Import another dataset
#general_countries <- read.csv('gender_development.csv')

#Rename columns names
colnames(general_countries)[3] <-"GENDER_DEVELOPMENT"
colnames(general_countries)[6] <-"FEMALE_LIFE_EXPENTANCY_ON_BIRTH"
colnames(general_countries)[7] <-"MALE_LIFE_EXPENTANCY_ON_BIRTH"
colnames(general_countries)[8] <-"FEMALE_YEARS_LEARNING"
colnames(general_countries)[9] <-"MALE_YEARS_LEARNING"
colnames(general_countries)[12] <-"GDP"

#Change some columns to be numeric
general_countries$GENDER_DEVELOPMENT = as.numeric(as.character(general_countries$GENDER_DEVELOPMENT))
general_countries$ FEMALE_LIFE_EXPENTANCY_ON_BIRTH=as.numeric(as.character(general_countries$FEMALE_LIFE_EXPENTANCY_ON_BIRTH))
general_countries$ MALE_LIFE_EXPENTANCY_ON_BIRTH= as.numeric(as.character(general_countries$MALE_LIFE_EXPENTANCY_ON_BIRTH))
general_countries$FEMALE_YEARS_LEARNING = as.numeric(as.character(general_countries$FEMALE_YEARS_LEARNING))
general_countries$MALE_YEARS_LEARNING = as.numeric(as.character(general_countries$MALE_YEARS_LEARNING))
general_countries$GDP = as.numeric(as.character(general_countries$GDP))

#Display the head
head(general_countries)
```


```{r}
#Get the difference of expected education among males and females
general_countries$EXPECTED_EDUCATION_YEARS = (general_countries$FEMALE_YEARS_LEARNING + general_countries$MALE_YEARS_LEARNING)/2

##Get the difference of expected life on birth among males and females

general_countries$EXPECTED_lIFE_ON_BIRTH = (general_countries$FEMALE_LIFE_EXPENTANCY_ON_BIRTH + general_countries$MALE_LIFE_EXPENTANCY_ON_BIRTH )/2

```


```{r}
#Add and match GDP, EXPECTED EDUCATION YEARS, EXPECTED LIFE ON BIRTH to the COUNTRIES_MEANS data.

Countries_means$GDP = general_countries$GDP[match(Countries_means$Country, general_countries$Country)]

Countries_means$EXPECTED_EDUCATION_YEARS<-general_countries$EXPECTED_EDUCATION_YEARS[match(Countries_means$Country, Countries_means$Country)]

Countries_means$EXPECTED_lIFE_ON_BIRTH<-general_countries$EXPECTED_lIFE_ON_BIRTH[match(Countries_means$Country, general_countries$Country)]

#Display the head
head(Countries_means)
```

```{r}
str(Countries_means)
```

### Correlation

To calculate the relationship between the continuous variables, we have used the spearman correlation as the data is not normally distributed which is important condition to use the standard correlation (pearson).


```{r}
cor(Countries_means$ConfirmedCases, Countries_means$GDP, method = "spearman", use = "complete.obs")
```

The Correlation between the GDP (Growth Development Progress) and the infection is : 0.5531839 (medium positive relationship). GDP is an economical definition but I find it to be indication that rich countries have specific features that participate in the spreading of the COVID-19 like : food, education, etc.

```{r}
cor(Countries_means$ConfirmedCases, Countries_means$EXPECTED_lIFE_ON_BIRTH, method = "spearman", use = "complete.obs")

```

Correlation between the EXPECTED_lIFE_ON_BIRTH and the infection is : 0.5781181 (also medium positive relationship).
We can think of the feature EXPECTED_lIFE_ON_BIRTH is an Indecation to the quality of healthcare system in the country, eventhough it has a medium relationship with the spreading.

```{r}
cor(Countries_means$EXPECTED_EDUCATION_YEARS, Countries_means$ConfirmedCases, method = "spearman", use = "complete.obs")

```

The correlation between the EXPECTED_EDUCATION_YEARS and the infection is : -0.09366682 (weak negative relationship).
 
```{r}
#Import another data that has information the gender and age of COVID19 patients
data2 = read.csv('metadata.csv')

#Remve unnecassary features

data2 = data2[c(2,3,4,5)]
head(data2)

```

```{r}
data2<-data2[(data2$finding=="COVID-19"),]

```


```{r}
#Show the distribution of COVID-19 among males and females
barplot(prop.table(table(data2$sex)))
```

 

This barplot shows that males have higher infection rate than females.  


```{r}
hist(data2$age)
```

People between 40 to 70 years of age are more infected with COVID-19. The highest infection rate is in the age range of 40 to 70 years old, Which makes sense as the immunity becomes weaker as we advance in age.


```{r}

#more graph to show the relationship between age and gender
ggplot(data2, 
       aes(x = age, 
           fill = sex)) + 
  geom_histogram(position = "stack")
```

```{r}
#Import another data that has useful information about the usage of fat and dairy supplement all over the countries.
data3 = read.csv('Fat_Supply_Quantity_Data.csv')
```

```{r}
#Select important columns only
data3 = data3[c(1,3,4,6,7,8,9,10,11,12,13)]
#Display the data
head(data3)
```



```{r}
#Rename columns names:

colnames(data3)[2] <-"animal_products"
colnames(data3)[3] <-"animal_fats"
colnames(data3)[4] <-"Cerial_excluding_beer"
colnames(data3)[5] <-"Eggs"
colnames(data3)[6] <-"Fish"
colnames(data3)[7] <-"Fruits"
colnames(data3)[8] <-"Meat"
colnames(data3)[9] <-"Miscellaneous"
colnames(data3)[10] <-"Milk"

#Match and add these columns to our previous data (Countries_means)

Countries_means$animal_products <- data3$animal_products[match(Countries_means$Country, data3$Country)]
Countries_means$animal_fats <- data3$animal_fats[match(Countries_means$Country, data3$Country)]
Countries_means$Cerial_excluding_beer <- data3$Cerial_excluding_beer[match(Countries_means$Country, data3$Country)]
Countries_means$Eggs <- data3$Eggs[match(Countries_means$Country, data3$Country)]
Countries_means$Fish <- data3$Fish[match(Countries_means$Country, data3$Country)]
Countries_means$Fruits <- data3$Fruits[match(Countries_means$Country, data3$Country)]
Countries_means$Meat <- data3$Meat[match(Countries_means$Country, data3$Country)]
Countries_means$Miscellaneous <- data3$Miscellaneous[match(Countries_means$Country, data3$Country)]
Countries_means$Milk <- data3$Milk[match(Countries_means$Country, data3$Country)]

print(Countries_means)
```



```{r}
#data4 = read.csv('countries_worldGDP.csv')

print(data4)
```


```{r}

str(data4)

```


```{r}
#Change the Country column in Data4 and Countries_means to character.

Countries_means$Country <- as.character(Countries_means$Country)

data4$Country <- as.character(data4$Country)

#Rename column 5

colnames(data4)[5] <-"population_density"

#Match and merge population_desity column to countries_means data
Countries_means$population_density<- data4$population_density[match(Countries_means$Country, data4$Country)]


data4$Country[2] == Countries_means$Country[2]

print(Countries_means)

```

```{r}

#I have combined these datasets (data4 and Countries_mean) using the countries names, one of the datasets has misspelled in the countries names so I built a function to fixand match the country name.

i=1
for (c1 in Countries_means$Country){
  
  c_chars = strsplit(c1,'')
  c_chars=c_chars[[1]]
  true_numbers =  list()

    for (c2 in data4$Country){
      c2_chars = strsplit(c2,'')
      c2_chars=c2_chars[[1]]
    
      
    if (length(c_chars) >= length(c2_chars)-2){
        if  (((c_chars[1] == c2_chars[1]) & (c_chars[2] == c2_chars[2]) ) || ((c_chars[3] == c2_chars[3]) & (c_chars[2] == c2_chars[2]))){
      counter = 0
      for (char in c2_chars){
          
          setuation =is.null(c_chars[char])
          if (setuation == FALSE ){
            counter = counter+1
          }
      }
      
        }
        
      }
    true_numbers[[i]] <- counter
    replacement_index = match(c(max(unlist(true_numbers))),true_numbers)
    
    }
    
    replacement_word = data4$Country[replacement_index]
    Countries_means$Country[i]= replacement_word
    i=i+1
    }

```



```{r}

#Match and merge columns from data4 to Countries_mean data

Countries_means$population_density<- data4$population_density[match(Countries_means$Country, data4$Country)]
Countries_means$population_density <- gsub(',', '.', Countries_means$population_density)

Countries_means$Literacy<- data4$Literacy[match(Countries_means$Country, data4$Country)]
Countries_means$Literacy <- gsub(',', '.', Countries_means$Literacy)

Countries_means$Climate<- data4$Climate[match(Countries_means$Country, data4$Country)]
Countries_means$Climate <- gsub(',', '.', Countries_means$Climate)

Countries_means$Industry<- data4$Industry[match(Countries_means$Country, data4$Country)]
Countries_means$Industry <- gsub(',', '.', Countries_means$Industry)

Countries_means$Agriculture<- data4$Agriculture[match(Countries_means$Country, data4$Country)]
Countries_means$Agriculture <- gsub(',', '.', Countries_means$Agriculture)


#Change the Literacy, population_density, Climate, Industry, and Agriculture columns to be numeric
Countries_means$Literacy = as.numeric(as.character(Countries_means$Literacy))
Countries_means$population_density = as.numeric(as.character(Countries_means$population_density))
Countries_means$Climate = as.numeric(as.character(Countries_means$Climate))
Countries_means$Industry = as.numeric(as.character(Countries_means$Industry))
Countries_means$Agriculture = as.numeric(as.character(Countries_means$Agriculture))


#Rename some columns
colnames(data4)[10] <-"Literacy"
colnames(data4)[15] <-"Climate"
colnames(data4)[19] <-"Industry"
colnames(data4)[18] <-"Agriculture"

head(Countries_means)
```


```{r}
i=1
for ( c in Countries_means ){
  col_name= names(Countries_means[i])
  if (col_name == "Country"){
    print('text')
  }else{
  plot(data = Countries_means, x=c, y=Countries_means$mean_infection, ,
   xlab=col_name, ylab="mean_infection", pch=19)
  
  } 
  
  i=i+1
}
```

```{r}
corelations = data.frame(cor(Countries_means[,2:21], use = "complete.obs"))
corelations

```


```{r}
write.table(corrlations, "D:/mydata.csv", sep=",")


```


```{r}
aggregate(Countries_means$mean_infection,list(Countries_means$Climate),sum)

```

```{r}

Countries_means = Countries_means[-c(2,3,12,13,14,17,19,20,21)]

```

```{r}
NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
Countries_means[2:12]=replace(Countries_means[2:12], TRUE, sapply(Countries_means[2:12], NA2mean))
```


```{r}
sum(is.na(Countries_means))
```


```{r}
#final_sub1 = read.csv('Counties geo.csv')
final_sub1$population_per_k = final_sub1$Population / final_sub1$Area.in.km²
final_sub1 = final_sub1[-c(2,3)]

```

```{r}
#final_sub2 = read.csv('gender_development.csv')
#Rename columns names
colnames(final_sub2)[6] <-"FEMALE_LIFE_EXPENTANCY_ON_BIRTH"
colnames(final_sub2)[7] <-"MALE_LIFE_EXPENTANCY_ON_BIRTH"
colnames(final_sub2)[8] <-"FEMALE_YEARS_LEARNING"
colnames(final_sub2)[9] <-"MALE_YEARS_LEARNING"
colnames(final_sub2)[12] <-"GDP_female"
colnames(final_sub2)[13] <-"GDP_male"


#Change some columns to be numeric

final_sub2$ FEMALE_LIFE_EXPENTANCY_ON_BIRTH=as.numeric(as.character(final_sub2$FEMALE_LIFE_EXPENTANCY_ON_BIRTH))
final_sub2$ MALE_LIFE_EXPENTANCY_ON_BIRTH=as.numeric(as.character(final_sub2$MALE_LIFE_EXPENTANCY_ON_BIRTH))
final_sub2$FEMALE_YEARS_LEARNING = as.numeric(as.character(final_sub2$FEMALE_YEARS_LEARNING))
final_sub2$MALE_YEARS_LEARNING = as.numeric(as.character(final_sub2$MALE_YEARS_LEARNING))
final_sub2$GDP_female = as.numeric(as.character(final_sub2$GDP_female))
final_sub2$GDP_male = as.numeric(as.character(final_sub2$GDP_male))



```






```{r}
#final_sub4 = read.csv('Fat_Supply_Quantity_Data.csv')
final_sub4 = final_sub4[c(1,3,4,6,7,11,12)]

colnames(final_sub4)[2] <-"animal_products"


colnames(final_sub4)[3] <-"animal_fats"


colnames(final_sub4)[4] <-"Cerial_excluding_beer"

colnames(final_sub4)[7] <-"Milk"


```

```{r}
final_sub5 = data4
final_sub6 =Country


```


```{r}
final_data = read.csv('COVID19_open_line_list.csv')
final_data = final_data[c(2,3,6,15)]


```




```{r}
names(final_sub1)
final_data$population_per_k <- final_sub1$population_per_k[match(final_data$country, final_sub1$Country)]
final_data$Continent <- final_sub1$Continent[match(final_data$country, final_sub1$Country)]



```


```{r}
names(final_sub2)
final_data$Life_expectancy <- final_sub2$FEMALE_LIFE_EXPENTANCY_ON_BIRTH[match(final_data$country, final_sub2$Country)]

final_data$Life_expectancy <- final_sub2$MALE_LIFE_EXPENTANCY_ON_BIRTH[match(final_data$country, final_sub2$Country)]

final_data$Life_expectancy <- final_sub2$FEMALE_YEARS_LEARNING[match(final_data$country, final_sub2$Country)]

final_data$Life_expectancy <- final_sub2$MALE_YEARS_LEARNING[match(final_data$country, final_sub2$Country)]

final_data$GDP <- final_sub2$GDP_female[match(final_data$country, final_sub2$Country)]

final_data$GDP <- final_sub2$GDP_male[match(final_data$country, final_sub2$Country)]


```


```{r}
names(final_sub3)
```


```{r}
names(final_sub4)


final_data$animal_products <- final_sub4$animal_products[match(final_data$country, final_sub4$Country)]


final_data$animal_fats <- final_sub4$animal_fats[match(final_data$country, final_sub4$Country)]


final_data$Cerial_excluding_beer <- final_sub4$Cerial_excluding_beer[match(final_data$country, final_sub4$Country)]


final_data$Eggs <- final_sub4$Eggs[match(final_data$country, final_sub4$Country)]


final_data$Miscellaneous <- final_sub4$Miscellaneous[match(final_data$country, final_sub4$Country)]


final_data$Milk <- final_sub4$Milk[match(final_data$country, final_sub4$Country)]


```


```{r}

names(final_sub5)

final_data$Literacy <- final_sub4$Literacy[match(final_data$country, final_sub5$Country)]

final_data$Climate <- final_sub4$Climate[match(final_data$country, final_sub5$Country)]


```


# 6. Reseach 2: Machine learning and deep learning



```{r}
#Data source: https://www.covid19survivalcalculator.com/en/download

#Import the data

#final_data2 = read.csv('master_dataset.csv', header = FALSE)

#Data source:https://www.kaggle.com/ahmedmd/corona-observations?select=convertcsvy.json

final_data2 = read.csv('master_dataset.csv', header = FALSE)
result = fromJSON(file ="convertcsvy.json")
final_data <- do.call(rbind, result)
json_data_frame <- as.data.frame(final_data)
final_data2 <- rbind(final_data2, json_data_frame)


#Display the fist few rows of the data
head(final_data2)

```


```{r}
#Remove un-important/irrelevant features

final_data2 = final_data2[-c(1,2,4,5,6,20,22,23,24,25,26,29,39,40,41,42,43)]

#Remove the last column
final_data2 = final_data2 [-c(27)]
```


```{r}
#any values with empty string as NA should be removed
final_data2 = subset(final_data2, V7!="" || V8!=""|| V9!="" || V10!="" || V11!="" || V12!="" || V13!="" || V14!="" || V15!="" || V16!="" || V17!="" || V18!="" || V19!="" || V21!="" || V27!="" || V30!="" || V31!="" || V32!="" || V33!="" || V34!="" || V35!="" || V36!="" || V37!="" || V38!="")

#Remove unknown observation in sex column
final_data2 = subset(final_data2, V12!="unknown")

```


```{r}
#Create a function to substitue the header
header_rename <- function(df) {
  names(df) <- as.character(unlist(df[1,]))
  df[-1,]
}
final_data2=header_rename(final_data2)

#Subsit the NA values in smoking column
final_data2 = subset(final_data2, smoking!="")

```


```{r}
#change all values to string
bob <- data.frame(lapply(final_data2, as.character), stringsAsFactors=FALSE)

# Divide the positive and negative cases
final_data2_p = final_data2[final_data2$covid19_positive =="1",]
final_data2_n = final_data2[final_data2$covid19_positive =="0",]

#Reduce the negative cases to 770. Negative cases are very large than positive cases which will cause biase in the data due to in quality in number of psotive and negative cases.
final_data2_n = final_data2_n[1:770,]

#combine the negative and positive cases together.
new_data <- rbind(final_data2_p, final_data2_n)

```

```{r}
#Plot histogram 

plotHistFunc <- function(x, na.rm = TRUE) {               
  nm <- names(x)
  for (i in seq_along(nm)) {
    print(ggplot(x,aes_string(x = nm[i],fill = "covid19_positive")) + geom_bar()) }
}

plotHistFunc(new_data)
```


```{r}
#change height, weight, and bmi to numberic. It's important to change it to character first before numberic to aviod mistakes.

new_data$height =  as.numeric(as.character(new_data$height))
new_data$weight =  as.numeric(as.character(new_data$weight))
new_data$bmi =  as.numeric(as.character(new_data$bmi))
new_data$covid19_positive <- factor(new_data$covid19_positive)

new_data$sex = as.factor(as.numeric(new_data$sex))
new_data$blood_type = as.factor(as.numeric(new_data$blood_type))
new_data$smoking = as.factor(as.numeric(new_data$smoking))
new_data$alcohol = as.factor(as.numeric(new_data$alcohol))
new_data$cannabis = as.factor(as.numeric(new_data$cannabis))
new_data$amphetamines = as.factor(as.numeric(new_data$amphetamines))
new_data$cocaine = as.factor(as.numeric(new_data$cocaine))
new_data$age = as.factor(as.numeric(new_data$age))

```


```{r}
#Scalling and dummy variables

new_data2 = new_data
new_data2$height = scale(new_data2$height)
new_data2$weight = scale(new_data2$weight)
new_data2$bmi = scale(new_data2$bmi)

#change some features to numberic before we do scalling
new_data2$cannabis = scale(as.numeric(new_data2$cannabis))
new_data2$amphetamines = scale(as.numeric(new_data2$amphetamines))
new_data2$cocaine = scale(as.numeric(new_data2$cocaine))
new_data2$mdma = scale(as.numeric(as.character(new_data2$mdma)))
new_data2$lsd = scale(as.numeric(as.character(new_data2$lsd)))


names(new_data)
```


```{r}
#Dummy variables
new_data2 <- fastDummies::dummy_cols(new_data2)
```


```{r}
#Remove categorical features after we did the dummy variables.

new_data2 = new_data2[-c(1,2,3,7,8,9,14,15,16,17,18,19,20,21,22,23,24,25,26)]

names(new_data2)
```

```{r}
#names(new_data2)

new_data3 = new_data2
new_data2= new_data3
new_data2 = new_data2 [-c(79)]
new_data2$covid19_positive_1   = as.factor(new_data2$covid19_positive_1 )
colnames(new_data2)[52]= "alcohol_01"



## set the seed to make your partition reproducible

smp_size <- floor(0.80 * nrow(new_data2))
set.seed(123)
train_ind <- sample(seq_len(nrow(new_data2)), size = smp_size)

train <- new_data2[train_ind, ]
test <- new_data2[-train_ind, ]
```

## Machine learning model using Random Forest

```{r}
#Random Forest Classifier

rf <- randomForest(covid19_positive_1 ~ .,
  data=train,
  importance=TRUE,
  na.action=na.roughfix
  )

#names(train)

y_pred = predict(rf ,test )
y_pred2 = as.factor(y_pred)

confusionMatrix(y_pred, test$covid19_positive)


#str(train)
```

## Deep learning model

```{r}
#Slpit the input and output to train and test. 
X_train= train [-c(79)]
X_test= test [-c(79)]
Y_train= train [c(79)]
Y_test= test [c(79)]
```

```{r}
#Deep learning require the output to be dummy variables. 

Y_train = to_categorical(Y_train,2)

Y_test = to_categorical(Y_test,2)
```


```{r}
# Deep learning require a sequence of layers
Deep_model= keras_model_sequential()
```


```{r}
#Add layers to our model. In each layer, we should include units (nodes of the exite layers) activation (helps in non-linearity) but the in the first layer, we should input shape that include the number of features.
Deep_model %>% 
  layer_dense(units= 50, activation = "relu", input_shape = 99) %>%
  layer_dense(units= 70, activation = "sigmoid") %>%
  layer_dense(units= 2, activation = "softmax") %>%
```


```{r}
#Structure the neural network(deep learning model)
Summary (Deep_model)
 
#Compling the model. Use optimizer adam and use accuracy to measure performance and use crossentropy to measure the error.

Deep_model %>% compile(loss = "categorical_crossentropy", optimizer = optimizer_adam(), metrics= c("accuracy"))
```


```{r}
#History: we used validation from the training set with .1 and number of iteration is 50 and divided the data to bathes of 42. Recording the operation in (hist)
hist = Deep_model %>% fit(X_train,Y_train, epochs=50, batch_size= 42, validation_split= .1)
```

```{r}
#Deep Learning evaluation 
Deep_model %>% evlauate (X_test, Y_test)
```


```{r}
#Deep Model 
Deep_model %>% predict_classes(X_test)

```

```{r}
plot(hist)
```

# 7. Resources:


https://www.kaggle.com/sandeepsonare/data-on-covid19-by-our-world-in-data?fbclid=IwAR1vfO6-30CfFkAYH2B8uUZi9JuSshDJWKaoeW0N6Yd3b2o2Te43n-Km0sg

https://www.kaggle.com/fernandol/countries-of-the-world?fbclid=IwAR3ZZ32gnC3b7nH4rnskx3Y1d5BFvqFSA1jp-W-xTNJcEgoE86DJetu_XGc

https://www.kaggle.com/nabeelsajid917/covid-19-x-ray-10000-images?fbclid=IwAR1814NzCczB4VdRMlSvCZG7_HQBj8ZWGn2YJqGg7LlLW0-s_bOZ-uuZmJg

https://www.kaggle.com/nabeelsajid917/covid-19-x-ray-10000-images?fbclid=IwAR1814NzCczB4VdRMlSvCZG7_HQBj8ZWGn2YJqGg7LlLW0-s_bOZ-uuZmJg#covid19.model

https://www.geonames.org/countries/

Source: https://www.kaggle.com/fernandol/countries-of-the-world?fbclid=IwAR2r9kQQHPlbl3TINuOHjMSYKpPRWta6IHU4LwgxCf9A7rWj6cq557hQYYM



